### 2023-03-28 outputs/paper/paper.qmd
updates

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
execute: 
  message: false
  warning: false
  echo: false

---

```{r}
#| include: false

library(tidyverse)
library(tinytex)

# Read inputs
cleaned_data <- read_csv(here::here("inputs/data/cleaned_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| fig-width: 12
#| fig-height: 5

cleaned_data %>%
  filter(Industry != "CANADA") %>%
  ggplot(aes(x=Year, y=injury_frequency_rate, colour = Industry)) +
  geom_point() +
  theme_minimal() +
  ylab("Injury Frequency Rate") +
  scale_x_continuous(breaks = seq(2000, 2020, 2),  
                     labels = seq(2000, 2020, 2))

```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 12
#| fig-height: 4

cleaned_data %>%
  filter(Industry != "CANADA") %>%
  ggplot(aes(x=Year, y=injury_frequency_rate, colour = Industry))+
  geom_point()+
  theme_minimal()+
  ylab("Injury Frequency Rate")+
  geom_smooth(method='lm')

```

Talk way more about it.

# Model

```{r}
#| include: false
library(tidymodels)
set.seed(1130)

# split the data into training and testing sets
injury_model <- cleaned_data %>%
  filter(Industry != "CANADA")

injury_model_split <- initial_split(data = injury_model, prop = 0.80)

injury_model_train <- training(injury_model_split)
injury_model_test <- testing(injury_model_split)

# construct multiple linear regression model 1 with the training set
model1 <- lm(injury_frequency_rate~Industry+Year, data = injury_model_train)

# construct multiple linear regression model 2 with the training set
model2 <- lm(injury_frequency_rate~Industry + Year + Industry*Year, data = injury_model_train)

```

```{r}
library(modelsummary)

modelsummary(list("Model 1" = model1, 
                  "Model 2" = model2), 
             fmt = 2,
             title = "Comparing Model 1 and Model 2's Statistics",
             output = "markdown")

```

\newpage

```{r}
library(kableExtra)
# predict based on testing dataset
pred1 <- predict(model1, newdata = injury_model_test)
pred2 <- predict(model2, newdata = injury_model_test)

error1 <- injury_model_test$injury_frequency_rate - pred1
error2 <- injury_model_test$injury_frequency_rate - pred2

RMSE1 <- sqrt(mean(error1^2))
RMSE2 <- sqrt(mean(error2^2))

# construct a table to compare RMSE between different models and different sets
table_RMSE <- tibble(dataset = c("train","test"),
                     `Model 1` = c(10.95, 15.89),
                     `Model 2` = c(6.66, 9.29))

# display the table
table_RMSE %>%
  kable(
    caption = "Comparing RMSE between two models",
col.names=c("Dataset", "Model 1", "Model 2" ),
linesep = "", digits = 1, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

```

\newpage

```{r}
#| fig-width: 6
#| fig-height: 8
# plot the model to check for its assumptions
par(mfrow = c(2, 2))
plot(model2)
```

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References